---
title: "Preliminary Explorations"
author: "Shaina Trevino"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("naniar")
library(tidyverse)
library(nhanesA)
library(here)
library(rio)
library(colorblindr)
library(ggridges)
library(ggridges)
library(ungeviz)
theme_set(theme_minimal())
```

## Tidying Data 

First steps:
* import data for all years
* join files
  * join data frames accross years
  * join data frames accross types
* write to csv and place in project for reproducability 
* write code to import data
* recode categorical variables with nhanes codebook

```{r import_data_initial, include = FALSE}
# PRELIMINARY DATA IMPORTING - No need to run anymore
 # requires install.packages('nhanesA')

#Before 2005, they used different drug use surveys

# DEMOGRAPHICS
#demo_01 <- nhanes('DEMO_B')
# demographics from 2005-2006 (release cycle 4)
#demo_05 <- nhanes('DEMO_D')

# demographics from 2007-2008 (release cycle 5)
#demo_07 <- nhanes('DEMO_E')

# demographics from 2009-2010 (release cycle 6)
#demo_09 <- nhanes('DEMO_F')

# demographics from 2011-2012 (release cycle 7)
#demo_11 <- nhanes('DEMO_G')

# demographics from 2013-2014 (release cycle 8)
#demo_13 <- nhanes('DEMO_H')

# demographics from 2015-2016 (release cycle 9)
#demo_15 <- nhanes('DEMO_I')

# SUBSTANCE USE
# SU from 2005-2006 (release cycle 4)
#su_05 <- nhanes('DUQ_D')

# SU from 2007-2008 (release cycle 5)
#su_07 <- nhanes('DUQ_E')

# SU from 2009-2010 (release cycle 6)
#su_09 <- nhanes('DUQ_F')

# SU from 2011-2012 (release cycle 7)
#su_11 <- nhanes('DUQ_G')

# SU from 2013-2014 (release cycle 8)
#su_13 <- nhanes('DUQ_H')

# SU from 2015-2016 (release cycle 9)
#su_15 <- nhanes('DUQ_I')

# Prescription usage from 15-16 and 01-02
#pu_15 <- nhanes('RXQ_RX_I')
#pu_01 <- nhanes('RXQ_RX_B')


```

```{r join_data, include = FALSE}
# participant id is unique across the years

## joining demographic data
#join <- full_join(demo_05, demo_07)
#join1 <- full_join(join, demo_09)
#join2 <- full_join(join1, demo_11)
#join3 <- full_join(join2, demo_13)
#join4 <- full_join(join3, demo_15)
#demo <- full_join(join4, demo_01)

# checking if id is unique
#demo %>% count(SEQN) %>% filter(n > 1)

# # joining SU data
# sj <- full_join(su_05, su_07)
# sj1 <- full_join(sj, su_09)
# sj2 <- full_join(sj1, su_11)
# sj3 <- full_join(sj2, su_13)
# su <-full_join(sj3, su_15)

# checking if id is unique
# su %>% count(SEQN) %>% filter(n > 1)

# joining PU data
#puj <- pu_01 %>% 
 # rename(RXDDRUG = RXD240B,
       #  RXDDAYS = RXD260, 
       #  RXDCOUNT = RXD295)
  
#pu <- full_join(pu_15, puj)
```


```{r export_data, include = FALSE}
# Example of exporting data into csv files
#write.csv(demo, "demo.csv")
#write.csv(su, "su.csv")
#write.csv(pu, "pu.csv")

```

```{r import_data}
demo <- import(here("data", "demo.csv"),
               setclass = "tbl_df")

su <- import(here("data", "su.csv"),
             setclass = "tbl_df")

pu <- import(here("data", "pu.csv"),
             setclass = "tbl_df")
```


```{r tidy}
# changing codes to years
t_demo <- demo %>% 
  janitor::clean_names() %>% 
  mutate(year = case_when(sddsrvyr == 2 ~ 2001,
                          sddsrvyr == 3 ~ 2003,
                          sddsrvyr == 4 ~ 2005,
                          sddsrvyr == 5 ~ 2007,
                          sddsrvyr == 6 ~ 2009,
                          sddsrvyr == 7 ~ 2011,
                          sddsrvyr == 8 ~ 2013,
                          sddsrvyr == 9 ~ 2015,
                          TRUE ~ NA_real_))

# creating labels for categorical variables
t_demo <- t_demo %>% 
  mutate(ethnicity = case_when(ridreth3 == 1 ~ "Hispanic", # combined to be more succinct
                               ridreth3 == 2 ~ "Hispanic", 
                               ridreth3 == 3 ~ "White",
                               ridreth3 == 4 ~ "Black",
                               ridreth3 == 6 ~ "Asian",
                               ridreth3 == 7 ~ "Other/Multiracial",
                               TRUE ~ NA_character_)) 

t_demo <- t_demo %>% 
  mutate(income = case_when(indhhin2 == 1 ~ "$0 - $4,999",
                            indhhin2 == 2 ~ "$5,000 to $9,999",
                            indhhin2 == 3 ~ "10,000 to $14,999",
                            indhhin2 == 4 ~ "$15,000 to $19,999",
                            indhhin2 == 5 ~ "$20,000 to $24,999",
                            indhhin2 == 6 ~ "$25,000 to $34,999",
                            indhhin2 == 7 ~ "$35,000 to $44,999",
                            indhhin2 == 8 ~ "$45,000 to $54,999",
                            indhhin2 == 9 ~ "$55,000 to $64,999",
                            indhhin2 == 10 ~ "$65,000 to $74,999",
                            indhhin2 == 11 ~ "$75,000 and Over",
                            indhhin2 == 12 ~ "Over $20,000",
                            indhhin2 == 13 ~ "Under $20,000",
                            indhhin2 == 14 ~ "$75,000 to $99,999",
                            indhhin2 == 15 ~ "$100,000 and over",
                            indhhin2 == 77 ~ "Refused",
                            indhhin2 == 99 ~ "Don't know",
                            TRUE ~ NA_character_))

t_demo <- t_demo %>% 
  mutate(education = case_when(dmdeduc2 == 1 ~ "No diploma/GED",
                               dmdeduc2 == 2 ~ "No diploma/GED",
                               dmdeduc2 == 3 ~ "High school graduate/GED",
                               dmdeduc2 == 4 ~ "Some college",
                               dmdeduc2 == 5 ~ "College graduate or above",
                               TRUE ~ NA_character_))

# Coding SU variables
su <- su %>% 
  janitor::clean_names() %>% 
  mutate(marijuana = case_when(duq200 == 1 ~ "Yes",
                               duq200 == 2 ~ "No",
                               TRUE ~ NA_character_)) %>% 
  mutate(cocaine = case_when(duq250 == 1 ~ "Yes",
                             duq250 == 2 ~ "No",
                             TRUE ~ NA_character_)) %>% 
  mutate(heroin = case_when(duq290 == 1 ~ "Yes",
                            duq290 == 2 ~ "No",
                            TRUE ~ NA_character_)) %>% 
  mutate(meth = case_when(duq330 == 1 ~ "Yes",
                          duq330 == 2 ~ "No",
                          TRUE ~ NA_character_))

# Coding sex variable
t_demo <- t_demo %>% 
   mutate(sex = case_when(riagendr == 1 ~ "Male", 
                         riagendr == 2 ~ "Female", 
                         TRUE ~ NA_character_)) 
# joining demo with su
d <- left_join(su, t_demo, by = "seqn")
```

### Initial Plot 1

General Audience
Goal: Show the trend in substance use from 2005 to 2015 by different substances and explore gender and racial differences
Variables: Substance use is the count of the number of people who reported use in that year

```{r initial_plot1}
h_count <- d %>% 
  group_by(year) %>% 
  count(heroin) %>% 
  filter(heroin == "Yes")

c_count <- d %>% 
  group_by(year) %>% 
  count(cocaine) %>% 
  filter(cocaine == "Yes")

me_count <- d %>% 
  group_by(year) %>% 
  count(meth) %>% 
  filter(meth == "Yes")

ma_count <- d %>% 
  group_by(year) %>% 
  count(marijuana) %>% 
  filter(marijuana == "Yes")

cj <- right_join(h_count, c_count, by = "year")
cj1 <- right_join(cj, me_count, by = "year")
su_count1 <- right_join(cj1, ma_count, by = "year")

su_count <- su_count1 %>% 
  ungroup() %>% 
  rename(Heroin = n.x,
         Cocaine = n.y,
         Meth = n.x.x,
         Marijuana = n.y.y) %>% 
  select(year, Heroin, Cocaine, Meth, Marijuana) %>% 
  gather(drug, n, Heroin:Marijuana) 

ggplot(su_count, aes(year, n)) + 
  geom_line(aes(color = drug))
```

Changes to make for revision:
* filter out marijuana to visualize illicit drugs per proposal
  * fct_reorder drug levels so that colors match order presented
  * annotate labels instead of legend - Since lines are so far apart, place label at end of trend line and delete the legend (will have to expand y limits)
* Change count to percent
  * increased cognitive load - small increases and harder to understand proportion
* edit x axis to show years more clearly
* make lines thicker, change color pallete
* Place points for the years in there is data for
* try facet by sex or ethnicity
  * stack vertically on top of each other (wont work with annotated labels)
  * change aspect ratio
* modify theme elements and place labels to reduce cognitive load
  * only major gridlines with data shown
  * Bigger facet titles
  * Adding labels
    * get rid of x axis since it is clear it is year
  * add space between y label and axis
  * center title
  * lighten axis text and caption
* Identify dips and rises in trends and search for reasons why (policy, etc)
  * Nov 2012 recreational marijuana legalized
  * Obama ends war on drugs in 2009
* Find best aspect ratio
* tidy data differently
  * get one dataset without joining
  * filter for data, gather, group_by, and count
* FIGURE OUT HOW TO PLACE ANNOTATES AT DIFFERENT POINTS ON FACETS AND GET WHITE BACKGROUND BEHIND ANNOTATE TEXT.

### Revised Plot 1

```{r revised_plot1_data}
rpd <- d %>% 
  filter(heroin == "Yes" |
         cocaine == "Yes" |
         meth == "Yes" |
         marijuana == "Yes") %>% 
  select(heroin, cocaine, meth, marijuana, year, sex, ethnicity) %>% 
  rename(Heroin = heroin,
         Cocaine = cocaine,
         Meth = meth,
         Marijuana = marijuana) %>% 
  gather(drug, response, Heroin:Marijuana) %>% 
  filter(response == "Yes") %>% 
  group_by(year, drug, sex) %>% 
  count() %>% 
  filter(drug != "Marijuana") %>% 
  ungroup() %>% 
  mutate(prop = (n/sum(n)))
```

```{r revised_plot1}
ggplot(rpd, aes(year, n, color = drug)) + 
  geom_line(size = 1.1, alpha = .7) +
  geom_point(size = 2.5) +
  facet_wrap(~ sex) +
  geom_text(data = filter(rpd, year == 2015),
            aes(y = n, label = drug),
            hjust = .4,
            vjust = -1.25,
            fontface = "bold") +
  geom_vline(xintercept = 2009.25, color = "gray30", linetype = "dotted", size = .6) +
  geom_vline(xintercept = 2012.8, color = "gray30", linetype = "dotted", size = .6) +
  annotate("text", label = "End of\n'War on Drugs'", x = 2007.7, y = 525, color = "gray30", size = 3) +
  annotate("text", label = "Marijuana 1st\nLeagalized", x = 2014.3, y = 525, color = "gray30", size = 3) + #nudging annotations for clarity
  scale_x_continuous(limits = c(2005, 2016),
                     breaks = c(2005, 2007, 2009, 2011, 2013, 2015),
                     labels = c(2005, "'07", "'09", "'11", "'13", 2015)) +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        panel.grid.major = element_line(color = "#e2dfd7"),
        panel.grid.minor = element_line(color = "#e2dfd7"),
        axis.title.x = element_blank(),
        panel.background = element_rect(fill = "#f9f6ef", color = "#f9f6ef"), 
        plot.background = element_rect(fill = "#f9f6ef", color = "#f9f6ef"),
        strip.text = element_text(color = "black", face = "bold", size = 12),
        plot.title = element_text(size = 14, hjust = .25),
        plot.caption = element_text(color = "gray50", size = 7),
        axis.text = element_text(color = "gray50")) +
  labs(y = "People reporting substance use\n",
       title = "Substance use trend from 2005 to 2015 in the U.S.",
       caption = "Data from the CDC's National Health and Nutrition Examination Survey")
  scale_color_OkabeIto() 
```


### Initial Plot 2

Scientific Audience
Goal: Show the change in distribution of how many days a person used different addictive benzodiazepines from year 2001 to 2015
Variables: days of use is continuous and types of benzodiazepines and year is categorical

```{r plot2_data}
pud1 <- pu %>% 
  janitor::clean_names() %>% 
  filter(rxddrug == "LORAZEPAM" |
         rxddrug == "ALPRAZOLAM" |
         rxddrug == "DIAZEPAM" |
         rxddrug == "CLONAZEPAM") %>% 
  select(seqn, rxddrug, rxddays, rxdcount) %>% 
  rename(drug = rxddrug,
         days = rxddays, 
         count = rxdcount)

to_join <- t_demo %>% 
  rename(age = ridageyr) %>% 
  select(seqn, year, age, sex) 

pud <- left_join(pud1, to_join, by = "seqn") %>% 
  naniar::replace_with_na(replace = list(days = 99999)) %>% 
  mutate(year = factor(year))
  
```


```{r initial_plot2}
ggplot(pud, aes(drug, days)) +
  geom_jitter(aes(color = year))
```

To refine:
* change to ridgeline density to clearly see them since so much overlap
* add variability with jitter
  * added jittered points
* change drug names to generic names for non-medical scientists 
  * change case of drug categories
* change days to years 
  * limit x axis to 0 on ggridges (never less than 0 years)
* change scale of ridges to to get less overlap and see jittered points better
* professional color pallete
* get rid of legend if adding color
* limit data to those under 80
* extend aspect ratio - wider for the years
* add facet boxes
* fct reorder density plots

```{r refined_plot2_NOTUSING}
rpud <- pud %>% 
  mutate(use_yr = days/365) %>% 
  filter(age != 80)

ggplot(rpud, aes(use_yr, fct_reorder(drug, use_yr), fill = drug)) +
  geom_density_ridges(aes(color = drug),
                      scale = .95,
                      alpha = .5, 
                      jittered_points = TRUE, 
                      position = position_points_jitter(width = .1, 
                                                        height = 0), 
                      point_shape = '|', 
                      point_size = 2, 
                      point_alpha = .4) +
  facet_wrap(~ year) +
  scale_y_discrete(breaks = c("LORAZEPAM", "CLONAZEPAM", "DIAZEPAM", "ALPRAZOLAM"),
                   labels = c("Ativan", "Klonopin", "Valium", "Xanax")) + # or delete and change case str_to_title and use specific name?
  scale_x_continuous(limits = c(0,40)) +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        strip.text = element_text(color = "gray30", face = "bold", size = 12),
        plot.caption = element_text(color = "gray50", size = 7)) +
  labs(x = "Years of use",
       title = "Distribution of Addictive Benzodiazepine Prescription Use",
       subtitle = "Differences in 2001 vs 2015 use in the U.S.",
       caption = "Data from the CDC's National Health and Nutrition Examination Survey") +
  scale_fill_viridis_d() +
  scale_color_viridis_d()
```

2ND Revision:
* change color to white to not get rideline where there is no data (0 on x)
* change facet to drug to see differences in years more clearly
  * facet by drug type and put year on y axis, so it is easier to see differences between the years over the drug types
* figure out how to change point color to black to see actual distribution of points

```{r refined_plot2}
rpud <- pud %>% 
  mutate(use_yr = days/365,
         drug = case_when(drug == "LORAZEPAM" ~ "Ativan",
                          drug == "CLONAZEPAM" ~ "Klonopin",
                          drug == "DIAZEPAM" ~ "Valium",
                          drug == "ALPRAZOLAM" ~ "Xanax",
                          TRUE ~ NA_character_)) %>% 
  filter(age != 80)

ggplot(rpud, aes(use_yr, year, fill = drug)) +
  geom_density_ridges(color = "white",
                      scale = .9,
                      alpha = .5, 
                      jittered_points = TRUE, 
                      position = position_points_jitter(width = .1, 
                                                        height = 0), 
                      point_shape = '|', 
                      point_size = 2, 
                      point_alpha = .35,
                      point_color = "black") +
  facet_wrap(~ drug) +
  scale_x_continuous(limits = c(0,40)) +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_text(face = "bold"),
        axis.text.x = element_text(color = "gray50"),
        strip.text = element_text(color = "gray30", face = "bold", size = 12),
        plot.caption = element_text(color = "gray50", size = 7)) +
  labs(x = "Years of use",
       title = "Distribution of Addictive Benzodiazepine Prescription Use",
       subtitle = "Differences in 2001 vs 2015 use in the U.S.",
       caption = "Data from the CDC's National Health and Nutrition Examination Survey\n2001 Sample (n = 18,418) and 2015 Sample (n = 19,647)") +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

#aspect ratio preview not working
```


### Initial Plot 3

Audience: Scientific community
Goal: Show how age of first substance use differs by income for each substance type
Variables: age of first substance use is categorical, income and substance types are categorical

```{r initial_plot 3}
pd <- d %>% 
  rename(age_Marijuana = duq210, 
         age_Cocaine = duq260, 
         age_Heroin = duq300, 
         age_Meth = duq340) %>% 
  filter(year != 2005) %>% 
  select(age_Cocaine, age_Marijuana, age_Heroin, age_Meth, income, sex, ethnicity, sex) %>% 
  gather(drug, age_use, age_Cocaine:age_Meth) %>% 
  separate(drug, c(NA, "drug"), sep = "_") %>% 
  drop_na(age_use)

ggplot(pd, aes(income, age_use)) + 
  geom_col(aes(fill = sex), position = "dodge") + #barely any sex differences
  facet_wrap(~ drug)
  
```

For Revision:
* take out sex since there are barely any differences
* create one average value to plot instead of distribution
* condense income levels
* have drug type on x axis and facet by income level (see crimes in denver hw2code)
* change colors and theme layout
* think of other type of graph (have point estimate with st. error since calculating average)
* fct_reorder levels
* take out marijana? illicit use
* filter out and combine income levels
* relevel income levels from lowest to highest
* filter out age of use that are coded as missing or denied
* calculate mean to plot instead of whatever is plotted above
* make error bars for uncertainty
  * calculate se for mean of age of use

```{r revised_plot3}
pd1 <- d %>% 
  rename(age_Marijuana = duq210, 
         age_Cocaine = duq260, 
         age_Heroin = duq300, 
         age_Meth = duq340) %>% 
  filter(year != 2005) %>% 
  select(age_Cocaine, age_Marijuana, age_Heroin, age_Meth, income, sex, year, ethnicity, education) %>% 
  gather(drug, age_use, age_Cocaine:age_Meth) %>% 
  separate(drug, c(NA, "drug"), sep = "_") %>% 
  mutate(income = case_when(income == "$0 - $4,999" ~ "Under $20,000", #grouped by 10s instead of 5s
                            income == "$5,000 to $9,999" ~ "Under $20,000",
                            income == "10,000 to $14,999" ~ "Under $20,000",
                            income == "$15,000 to $19,999" ~ "Under $20,000",
                            income == "$20,000 to $24,999" ~ "$20,000 to $34,999",
                            income == "$25,000 to $34,999" ~ "$20,000 to $34,999",
                            income == "$35,000 to $44,999" ~ "$35,000 to $54,999",
                            income == "$45,000 to $54,999" ~ "$35,000 to $54,999",
                            income == "$55,000 to $64,999" ~ "$55,000 to $99,999",
                            income ==  "$65,000 to $74,999" ~ "$55,000 to $99,999",
                            income ==  "$75,000 and Over" ~ "$75,000 and Over",
                            income ==  "Over $20,000" ~ "Over $20,000",
                            income ==  "Under $20,000" ~ "Under $20,000",
                            income ==  "$75,000 to $99,999" ~ "$55,000 to $99,999",
                            income ==  "$100,000 and over" ~ "$100,000 and over",
                            TRUE ~ NA_character_),
         income = factor(income,
                         levels = c("Under $20,000",
                                    "$20,000 to $34,999", 
                                    "$35,000 to $54,999",
                                    "55,000 to $99,999",
                                    "$100,000 and over"))) %>% 
  naniar::replace_with_na(replace = list(age_use = c(999, 777, 99, 0, 77))) %>% 
  drop_na(age_use)
          
sum_pd1 <- pd1 %>% 
  filter(income != "$75,000 and Over" |
         income != "Over $20,000") %>% 
  group_by(drug, income) %>% 
  summarize(avg_age = mean(age_use),
            n = length(age_use),
            sd = sd(age_use),
            se = sd / sqrt(n))


ggplot(sum_pd1, aes(fct_reorder(drug, avg_age), avg_age)) + 
  geom_col() +
  facet_wrap(~ income)
#can't really see differences - make jitter with point estimate and se bar
#consider filtering for some years and taking out marijuana again

```

2nd revise:

* plot jittered raw distribution with point and uncertaintly over it
* plot using only 2 years recent data for less overlap/confounds
* add label for average age
* could try stat density

```{r revised_data_p3}
pd1_rev <- pd1 %>% 
  mutate(education = factor(education,
                            levels = c("College graduate or above",
                                       "Some college", 
                                       "High school graduate/GED",
                                       "No diploma/GED",
                                       TRUE ~ NA_character_)),
         ethnicity_group = case_when(ethnicity == "Hispanic" ~ "Non- White", # combined to be more succinct
                                     ethnicity == "Hispanic" ~ "Non- White", 
                                     ethnicity == "White" ~ "White",
                                     ethnicity == "Black" ~ "Non- White",
                                     ethnicity == "Asian" ~ "Non- White",
                                     TRUE ~ NA_character_),
         drug = factor(drug,
                       levels = c("Marijuana",
                                  "Cocaine",
                                  "Meth",
                                  "Heroin"))) %>% 
  drop_na(age_use, education)
```


```{r revised_plot2}
pd1_rev_filter <- pd1_rev %>% 
  filter(year == 2015 | 
         year == 2013 ) %>% 
  filter(drug != "Marijuana")

pd1_rev_sum <- pd1_rev %>%  
  filter(year == 2015 |
         year == 2013 ) %>% 
  group_by(drug, education) %>% 
  summarize(avg_age = mean(age_use),
            n = length(age_use),
            sd = sd(age_use),
            se = sd / sqrt(n)) %>% 
  drop_na(education) %>% 
  filter(drug != "Marijuana")

p <- ggplot(pd1_rev_sum, aes(education, avg_age)) + 
  geom_jitter(aes(x = education, y = age_use, color = education), data = pd1_rev_filter, alpha = .1, width = .25) +
  geom_errorbar(aes(ymin = avg_age + qnorm(.025)*se,
                    ymax = avg_age + qnorm(.975)*se),
                width = .4) +
  geom_point(size = 1) + # add jitter from other dataset (before point)
  facet_wrap(~ drug, ncol = 1) + # added ncol = 1 to see more of variability 
  coord_flip() + #coord flip to see labels better and differences in outcome
  theme(legend.position = "none",
        plot.caption = element_text(color = "gray50", size = 7)) +
  labs(y = "Age of first use (in years)",
       x = "",
       title = "Age of First Substance Use by Education Level",
       subtitle = "Distribution and point estimate for average age of first use",
       caption = "Data from the CDC's National Health and Nutrition Examination Survey (2013-2015)")
p
```
 
* trying out darker theme

```{r dark_theme}
p + theme(plot.background = element_rect(fill = "gray10"),
        axis.text = element_text(colour = "gray80"),
        axis.title = element_text(colour = "gray80"),
        plot.title = element_text(colour = "gray80"),
        plot.subtitle = element_text(colour = "gray80"),
        plot.caption = element_text(colour = "gray80", vjust = 1),
        panel.grid.major = element_line(colour = "gray30"), 
        panel.grid.minor = element_line(colour = "gray30")) +
  geom_errorbar(aes(ymin = avg_age + qnorm(.025)*se,
                    ymax = avg_age + qnorm(.975)*se),
                width = .4,
                color = "gray90") +
  geom_point(size = 1, color = "gray90")

```
 
* For a general audience
 
```{r density}
ggplot(pd1_rev_sum, aes(avg_age, education)) + 
  stat_confidence_density(aes(moe = se),
                          fill = "cornflowerblue",
                          height = .45) +
  xlim(15, 35) +
  geom_point(alpha = .8, size = 1.25) +
  facet_wrap(~drug, ncol = 1) +
  geom_text(aes(avg_age, education, label = round(avg_age, 1)),
            nudge_x = 1.5,
            nudge_y = .4,
            size = 3,
            color = "gray70") +
  labs(title = "Differences in average age of first substance use based on education level",
       x = "Average age of first use (in years)",
       y = "") +
  theme(plot.background = element_rect(fill = "gray10"),
        axis.text = element_text(colour = "gray80"),
        axis.title = element_text(colour = "gray80"),
        plot.title = element_text(colour = "gray80", hjust = 1.5),
        plot.caption = element_text(colour = "gray80", vjust = 1),
        panel.grid.major = element_line(colour = "gray30"), 
        panel.grid.minor = element_line(colour = "gray30"),
        strip.text = element_text(color = "gray80"))

#limit x axis
# add labels for average age since limiting x axis
  
```

* exploring ethnicity differences - initial (most likley not using)

```{r ethnicity_diff_plot3}
pd_sum_eth <- pd1_rev %>% 
  filter(year == 2015) %>% 
  group_by(drug, education, ethnicity_group) %>% 
  summarize(avg_age = mean(age_use),
            n = length(age_use),
            sd = sd(age_use),
            se = sd / sqrt(n)) %>% 
  drop_na(education, ethnicity_group)

ggplot(pd_sum_eth, aes(education, avg_age)) +
  geom_col(aes(fill = fct_reorder(ethnicity_group, avg_age)), position = "dodge") +
  facet_wrap(~ fct_reorder(drug, avg_age), nrow = 1)

ggplot(pd_sum_eth, aes(education, avg_age)) +
  geom_col(aes(fill = fct_reorder(ethnicity_group, avg_age)), position = "dodge") +
  facet_wrap(~ fct_reorder(drug, avg_age), ncol = 1) +
  coord_flip()

```


