---
title: "Preliminary Explorations"
author: "Shaina Trevino"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# install.packages("nhanesA")
# install.packages("naniar")
library(nhanesA)
library(here)
library(rio)
theme_set(theme_minimal())
```
## Tidying Data 

import data
join files
write to csv

```{r import_data}

#Before 2005, they used different drug use surveys

# DEMOGRAPHICS
#demo_01 <- nhanes('DEMO_B')
# demographics from 2005-2006 (release cycle 4)
#demo_05 <- nhanes('DEMO_D')

# demographics from 2007-2008 (release cycle 5)
#demo_07 <- nhanes('DEMO_E')

# demographics from 2009-2010 (release cycle 6)
#demo_09 <- nhanes('DEMO_F')

# demographics from 2011-2012 (release cycle 7)
#demo_11 <- nhanes('DEMO_G')

# demographics from 2013-2014 (release cycle 8)
#demo_13 <- nhanes('DEMO_H')

# demographics from 2015-2016 (release cycle 9)
#demo_15 <- nhanes('DEMO_I')

# SUBSTANCE USE
# SU from 2005-2006 (release cycle 4)
#su_05 <- nhanes('DUQ_D')

# SU from 2007-2008 (release cycle 5)
#su_07 <- nhanes('DUQ_E')

# SU from 2009-2010 (release cycle 6)
#su_09 <- nhanes('DUQ_F')

# SU from 2011-2012 (release cycle 7)
#su_11 <- nhanes('DUQ_G')

# SU from 2013-2014 (release cycle 8)
#su_13 <- nhanes('DUQ_H')

# SU from 2015-2016 (release cycle 9)
#su_15 <- nhanes('DUQ_I')

# Prescription usage from 15-16 and 01-02
#pu_15 <- nhanes('RXQ_RX_I')
#pu_01 <- nhanes('RXQ_RX_B')


```

```{r join_data}
# participant id is unique across the years

## joining demographic data
#join <- full_join(demo_05, demo_07)
#join1 <- full_join(join, demo_09)
#join2 <- full_join(join1, demo_11)
#join3 <- full_join(join2, demo_13)
#join4 <- full_join(join3, demo_15)
#demo <- full_join(join4, demo_01)

# checking if id is unique
#demo %>% count(SEQN) %>% filter(n > 1)

# # joining SU data
# sj <- full_join(su_05, su_07)
# sj1 <- full_join(sj, su_09)
# sj2 <- full_join(sj1, su_11)
# sj3 <- full_join(sj2, su_13)
# su <-full_join(sj3, su_15)

# checking if id is unique
# su %>% count(SEQN) %>% filter(n > 1)

# joining PU data
#puj <- pu_01 %>% 
 # rename(RXDDRUG = RXD240B,
       #  RXDDAYS = RXD260, 
       #  RXDCOUNT = RXD295)
  
#pu <- full_join(pu_15, puj)
```


```{r export_data}
# Example of exporting data into csv files
#write.csv(demo, "demo.csv")
#write.csv(su, "su.csv")
#write.csv(pu, "pu.csv")

```

```{r import_data}
demo <- import(here("data", "demo.csv"),
               setclass = "tbl_df")

su <- import(here("data", "su.csv"),
             setclass = "tbl_df")

pu <- import(here("data", "pu.csv"),
             setclass = "tbl_df")
```


```{r tidy}
# changing codes to years
t_demo <- demo %>% 
  janitor::clean_names() %>% 
  mutate(year = case_when(sddsrvyr == 2 ~ 2001,
                          sddsrvyr == 3 ~ 2003,
                          sddsrvyr == 4 ~ 2005,
                          sddsrvyr == 5 ~ 2007,
                          sddsrvyr == 6 ~ 2009,
                          sddsrvyr == 7 ~ 2011,
                          sddsrvyr == 8 ~ 2013,
                          sddsrvyr == 9 ~ 2015,
                          TRUE ~ NA_real_))

# creating labels for categorical variables
t_demo <- t_demo %>% 
  mutate(ethnicity = case_when(ridreth3 == 1 ~ "Hispanic", # combined to be more succinct
                               ridreth3 == 2 ~ "Hispanic", 
                               ridreth3 == 3 ~ "White",
                               ridreth3 == 4 ~ "Black",
                               ridreth3 == 6 ~ "Asian",
                               ridreth3 == 7 ~ "Other/Multiracial",
                               TRUE ~ NA_character_)) 

t_demo <- t_demo %>% 
  mutate(income = case_when(indhhin2 == 1 ~ "$0 - $4,999",
                            indhhin2 == 2 ~ "$5,000 to $9,999",
                            indhhin2 == 3 ~ "10,000 to $14,999",
                            indhhin2 == 4 ~ "$15,000 to $19,999",
                            indhhin2 == 5 ~ "$20,000 to $24,999",
                            indhhin2 == 6 ~ "$25,000 to $34,999",
                            indhhin2 == 7 ~ "$35,000 to $44,999",
                            indhhin2 == 8 ~ "$45,000 to $54,999",
                            indhhin2 == 9 ~ "$55,000 to $64,999",
                            indhhin2 == 10 ~ "$65,000 to $74,999",
                            indhhin2 == 11 ~ "$75,000 and Over",
                            indhhin2 == 12 ~ "Over $20,000",
                            indhhin2 == 13 ~ "Under $20,000",
                            indhhin2 == 14 ~ "$75,000 to $99,999",
                            indhhin2 == 15 ~ "$100,000 and over",
                            indhhin2 == 77 ~ "Refused",
                            indhhin2 == 99 ~ "Don't know",
                            TRUE ~ NA_character_))

# Coding SU variables
su <- su %>% 
  janitor::clean_names() %>% 
  mutate(marijuana = case_when(duq200 == 1 ~ "Yes",
                               duq200 == 2 ~ "No",
                               TRUE ~ NA_character_)) %>% 
  mutate(cocaine = case_when(duq250 == 1 ~ "Yes",
                             duq250 == 2 ~ "No",
                             TRUE ~ NA_character_)) %>% 
  mutate(heroin = case_when(duq290 == 1 ~ "Yes",
                            duq290 == 2 ~ "No",
                            TRUE ~ NA_character_)) %>% 
  mutate(meth = case_when(duq330 == 1 ~ "Yes",
                          duq330 == 2 ~ "No",
                          TRUE ~ NA_character_))

# Coding sex variable
t_demo <- t_demo %>% 
   mutate(sex = case_when(riagendr == 1 ~ "Male", 
                         riagendr == 2 ~ "Female", 
                         TRUE ~ NA_character_)) 
# joining demo with su
d <- left_join(su, t_demo, by = "seqn")
```

### Initial Plot 1

Goal: Show the trend in substance use from 2005 to 2015 by different substances and explore gender and racial differences
Variables: Substance use is the count of the number of people who reported use in that year

```{r initial_plot1}
h_count <- d %>% 
  group_by(year) %>% 
  count(heroin) %>% 
  filter(heroin == "Yes")

c_count <- d %>% 
  group_by(year) %>% 
  count(cocaine) %>% 
  filter(cocaine == "Yes")

me_count <- d %>% 
  group_by(year) %>% 
  count(meth) %>% 
  filter(meth == "Yes")

ma_count <- d %>% 
  group_by(year) %>% 
  count(marijuana) %>% 
  filter(marijuana == "Yes")

cj <- right_join(h_count, c_count, by = "year")
cj1 <- right_join(cj, me_count, by = "year")
su_count1 <- right_join(cj1, ma_count, by = "year")

su_count <- su_count1 %>% 
  ungroup() %>% 
  rename(Heroin = n.x,
         Cocaine = n.y,
         Meth = n.x.x,
         Marijuana = n.y.y) %>% 
  select(year, Heroin, Cocaine, Meth, Marijuana) %>% 
  gather(drug, n, Heroin:Marijuana) 

ggplot(su_count, aes(year, n)) + 
  geom_line(aes(color = drug))
```

Need to:
* take out marijuana (want to visualize illicit drugs)
* fct_reorder drug levels so that colors match order presented
* Identify dips and rises in trends and search for reasons why (policy, etc)
* change count to percent?
* make year a data variable
* make lines thicker, change color pallete
* modify theme elements and place labels to reduce cognitive load
* Since lines are so far apart, place label at end of trend line and delete the legend (will have to expand y limits)
* Place points for the years in there is data for (and then make line dotted?) and put label for high/low point years
* try facet by sex or ethnicity
* Nov 2012 recreational marijuana legalized
* Obama ends war on drugs in 2009
NEXT - explore differences related to race/ethnicity and gender (consider same plot for different audiences)

```{r revised_plot1_data}
rpd <- d %>% 
  filter(heroin == "Yes" |
         cocaine == "Yes" |
         meth == "Yes" |
         marijuana == "Yes") %>% 
  select(heroin, cocaine, meth, marijuana, year, sex, ethnicity) %>% 
  rename(Heroin = heroin,
         Cocaine = cocaine,
         Meth = meth,
         Marijuana = marijuana) %>% 
  gather(drug, response, Heroin:Marijuana) %>% 
  filter(response == "Yes") %>% 
  group_by(year, drug, sex) %>% 
  count() %>% 
  filter(drug != "Marijuana")

ggplot(rpd, aes(year, n)) + 
  geom_line(aes(color = drug), size = 1.1) +
  facet_wrap(~ sex)
```


### Initial Plot 2

Goal: Show the change in distribution of how many days a person used different addictive benzodiazepines from year 2001 to 2015
Variables: days of use is continuous and types of benzodiazepines and year is categorical

```{r plot2_data}
pud1 <- pu %>% 
  janitor::clean_names() %>% 
  filter(rxddrug == "LORAZEPAM" |
         rxddrug == "ALPRAZOLAM" |
         rxddrug == "DIAZEPAM" |
         rxddrug == "CLONAZEPAM") %>% 
  select(seqn, rxddrug, rxddays, rxdcount) %>% 
  rename(drug = rxddrug,
         days = rxddays, 
         count = rxdcount)

to_join <- t_demo %>% 
  rename(age = ridageyr) %>% 
  select(seqn, year, age, sex) 

pud <- left_join(pud1, to_join, by = "seqn") %>% 
  naniar::replace_with_na(replace = list(days = 99999)) %>% 
  mutate(year = factor(year))
  
```


```{r initial_plot2}
ggplot(pud, aes(drug, days)) +
  geom_jitter(aes(color = year))
```

To refine:
* change to ridgeline density to clearly see them since so much overlap
* facet by drug type and put year on y axis, so it is easier to see differences between the years over the drug types
* add variability? maybe with jitter
* change drug names to generic names?
* change days to years (limit x axis to 0 on ggridges)

```{r refined_plot2}
ggplot(pud, aes(days, drug)) +
  ggridges::geom_density_ridges(alpha = .6) +
  facet_wrap(~ year)
```

### Initial Plot 3

Goal: Show how age of first substance use differs by income for each substance type
Variables: age of first substance use is categorical, income and substance types are categorical

```{r initial_plot 3}
pd <- td %>% 
  rename(age_Marijuana = duq210, 
         age_Cocaine = duq260, 
         age_Heroin = duq300, 
         age_Meth = duq340) %>% 
  filter(year != 2005) %>% 
  select(age_Cocaine, age_Marijuana, age_Heroin, age_Meth, income, sex, ethnicity, sex) %>% 
  gather(drug, age_use, age_Cocaine:age_Meth) %>% 
  separate(drug, c(NA, "drug"), sep = "_") %>% 
  drop_na(age_use)

ggplot(pd, aes(income, age_use)) + 
  geom_col(aes(fill = sex), stat = "identity") +
  facet_wrap(~ drug)
  
```

For Revision:
* dodge bars
* create one average value to plot instead of distribution
* condense income levels
* have drug type on x axis and facet by income level (see crimes in denver hw2code)
* change colors and theme layout
* think of other type of graph (have point estimate with st. error since calculating average)
* fct_reorder levels
* take out marijana? illicit use
