---
title: "Preliminary Explorations"
author: "Shaina Trevino"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# install.packages("nhanesA")
# install.packages("naniar")
library(nhanesA)
library(here)
library(rio)
theme_set(theme_minimal())
```
## Tidying Data 

import data
join files
write to csv

```{r import_data}

#Before 2005, they used different drug use surveys

# DEMOGRAPHICS
# demographics from 2005-2006 (release cycle 4)
demo_05 <- nhanes('DEMO_D')

# demographics from 2007-2008 (release cycle 5)
demo_07 <- nhanes('DEMO_E')

# demographics from 2009-2010 (release cycle 6)
demo_09 <- nhanes('DEMO_F')

# demographics from 2011-2012 (release cycle 7)
demo_11 <- nhanes('DEMO_G')

# demographics from 2013-2014 (release cycle 8)
demo_13 <- nhanes('DEMO_H')

# demographics from 2015-2016 (release cycle 9)
demo_15 <- nhanes('DEMO_I')

# SUBSTANCE USE
# SU from 2005-2006 (release cycle 4)
su_05 <- nhanes('DUQ_D')

# SU from 2007-2008 (release cycle 5)
su_07 <- nhanes('DUQ_E')

# SU from 2009-2010 (release cycle 6)
su_09 <- nhanes('DUQ_F')

# SU from 2011-2012 (release cycle 7)
su_11 <- nhanes('DUQ_G')

# SU from 2013-2014 (release cycle 8)
su_13 <- nhanes('DUQ_H')

# SU from 2015-2016 (release cycle 9)
su_15 <- nhanes('DUQ_I')

# 

```

```{r join_data}
# participant id is unique across the years

# joining demographic data
join <- full_join(demo_05, demo_07)
join1 <- full_join(join, demo_09)
join2 <- full_join(join1, demo_11)
join3 <- full_join(join2, demo_13)
demo <- full_join(join3, demo_15)
# checking if id is unique
demo %>% count(SEQN) %>% filter(n > 1)

# joining SU data
sj <- full_join(su_05, su_07)
sj1 <- full_join(sj, su_09)
sj2 <- full_join(sj1, su_11)
sj3 <- full_join(sj2, su_13)
su <-full_join(sj3, su_15)
# checking if id is unique
su %>% count(SEQN) %>% filter(n > 1)

# joining demo with su
d <- left_join(su, demo)
```


```{r export_data}
# Example of exporting data into csv files
write.csv(demo, "demo.csv")
write.csv(su, "su.csv")

```

```{r tidy}
# changing codes to years
td <- d %>% 
  janitor::clean_names() %>% 
  mutate(year = case_when(sddsrvyr == 4 ~ 2005,
                          sddsrvyr == 5 ~ 2007,
                          sddsrvyr == 6 ~ 2009,
                          sddsrvyr == 7 ~ 2011,
                          sddsrvyr == 8 ~ 2013,
                          sddsrvyr == 9 ~ 2015,
                          TRUE ~ NA_real_))

# creating labels for categorical variables
td <- td %>% 
  mutate(ethnicity = case_when(ridreth3 == 1 ~ "Hispanic", # combined to be more succinct
                               ridreth3 == 2 ~ "Hispanic", 
                               ridreth3 == 3 ~ "White",
                               ridreth3 == 4 ~ "Black",
                               ridreth3 == 6 ~ "Asian",
                               ridreth3 == 7 ~ "Other/Multiracial",
                               TRUE ~ NA_character_)) 

td <- td %>% 
  mutate(income = case_when(indhhin2 == 1 ~ "$0 - $4,999",
                            indhhin2 == 2 ~ "$5,000 to $9,999",
                            indhhin2 == 3 ~ "10,000 to $14,999",
                            indhhin2 == 4 ~ "$15,000 to $19,999",
                            indhhin2 == 5 ~ "$20,000 to $24,999",
                            indhhin2 == 6 ~ "$25,000 to $34,999",
                            indhhin2 == 7 ~ "$35,000 to $44,999",
                            indhhin2 == 8 ~ "$45,000 to $54,999",
                            indhhin2 == 9 ~ "$55,000 to $64,999",
                            indhhin2 == 10 ~ "$65,000 to $74,999",
                            indhhin2 == 11 ~ "$75,000 and Over",
                            indhhin2 == 12 ~ "Over $20,000",
                            indhhin2 == 13 ~ "Under $20,000",
                            indhhin2 == 14 ~ "$75,000 to $99,999",
                            indhhin2 == 15 ~ "$100,000 and over",
                            indhhin2 == 77 ~ "Refused",
                            indhhin2 == 99 ~ "Don't know",
                            TRUE ~ NA_character_))

# Coding SU variables
td <- td %>% 
  mutate(marijuana = case_when(duq200 == 1 ~ "Yes",
                               duq200 == 2 ~ "No",
                               TRUE ~ NA_character_)) %>% 
  mutate(cocaine = case_when(duq250 == 1 ~ "Yes",
                             duq250 == 2 ~ "No",
                             TRUE ~ NA_character_)) %>% 
  mutate(heroin = case_when(duq290 == 1 ~ "Yes",
                            duq290 == 2 ~ "No",
                            TRUE ~ NA_character_)) %>% 
  mutate(meth = case_when(duq330 == 1 ~ "Yes",
                          duq330 == 2 ~ "No",
                          TRUE ~ NA_character_))
```


```{r initial_plot}
h_count <- td %>% 
  group_by(year) %>% 
  count(heroin) %>% 
  filter(heroin == "Yes")

c_count <- td %>% 
  group_by(year) %>% 
  count(cocaine) %>% 
  filter(cocaine == "Yes")

me_count <- td %>% 
  group_by(year) %>% 
  count(meth) %>% 
  filter(meth == "Yes")

ma_count <- td %>% 
  group_by(year) %>% 
  count(marijuana) %>% 
  filter(marijuana == "Yes")

cj <- right_join(h_count, c_count, by = "year")
cj1 <- right_join(cj, me_count, by = "year")
su_count1 <- right_join(cj1, ma_count, by = "year")

su_count <- su_count1 %>% 
  ungroup() %>% 
  rename(Heroin = n.x,
         Cocaine = n.y,
         Meth = n.x.x,
         Marijuana = n.y.y) %>% 
  select(year, Heroin, Cocaine, Meth, Marijuana) %>% 
  gather(drug, n, Heroin:Marijuana) 

ggplot(su_count, aes(year, n)) + 
  geom_line(aes(color = drug))
```

Need to:
* fct_reorder drug levels so that colors match order presented
* Identify dips and rises in trends and search for reasons why (policy, etc)
* change count to percent?
* make year a data variable
* make lines thicker, choose prettier color
* modify theme elements and place labels to reduce cognitive load
* Since lines are so far apart, place label at end of trend line and delete the legend (will have to expand y limits)
* Place points for the years in there is data for (and then make line dotted?)

NEXT - explore differences related to race/ethnicity and gender (consider same plot for different audiences)
